CUNY SPS DATA 605 Final

### 2019-05-21 DATA605/MSilva_Final.Rmd
More work on the final

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#output: html_document
#output:
#  rmdformats::html_clean:
#    highlight: kate
library(knitr)
#library(rmdformats)
library(dplyr)
library(ggplot2)
library(tidyr)
library(kableExtra)
```

@@

```{r, echo=TRUE, eval=TRUE}
kaggle <- read.csv("Kaggle_House_Prices/train.csv") %>%
  mutate(MSSubClass = as.factor(MSSubClass))
```

@@

    d_table <- kaggle %>%
      group_by(.dots = variable) %>%
      summarise(count = n()) %>%
      kable() %>%
      kable_styling()
  } else {
    p <- ggplot(kaggle, aes_string(variable)) + 
      geom_histogram(binwidth = 10)
    q <- quantile(d, na.rm = TRUE)
    d_table <- data.frame(Min = min(d, na.rm = TRUE),
                          First = q[2],
                          Median = median(d, na.rm = TRUE), 
                          Mean = mean(d, na.rm = TRUE),
                          Third = q[4],
                          Max = max(d, na.rm = TRUE)) %>%
      rename("First Q" = First,
             "Third Q" = Third) %>%
      gather(Measure, Value) %>%
      kable() %>%
      kable_styling()
  }
  print(p)
  print(d_table)
}
```

@@

```{r, echo=TRUE}
kaggle %>%
  select(TotalBsmtSF, GrLivArea, YearBuilt) %>%
  cor() %>%
  kable() %>%
  kable_styling()
```

Test the hypothesis that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval. Discuss the meaning of your analysis. Would you be worried about familywise error? Why or why not?

```{r}
d_table <- data.frame(Min = min(d, na.rm = TRUE),
                          "First Qu." = q[2],
                          Median = median(d, na.rm = TRUE), 
                          Mean = mean(d, na.rm = TRUE),
                          "Third Qu." = q[4],
                          Max = max(d, na.rm = TRUE)) %>%
  gather(Measure, Value)

```

@@

Invert your correlation matrix from above. (This is known as the precision matrix and contains variance inflation factors on the diagonal.) Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix. Conduct LU decomposition on the matrix.

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(MASS)
library(Matrix)
library(knitr)
library(matlib)
library(dplyr)
library(ggplot2)
library(tidyr)
library(kableExtra)
```

@@

```{r, echo=TRUE, eval=TRUE}
kaggle <- read.csv("Kaggle_House_Prices/train.csv") %>%
  filter(GrLivArea < 4000) %>% # Removing outliers per http://jse.amstat.org/v19n3/decock.pdf
  # Filling in missing values with zeros where it makes sense
  mutate(BedroomAbvGr = replace_na(BedroomAbvGr, 0),
         BsmtFullBath = replace_na(BsmtFullBath, 0),
         BsmtHalfBath = replace_na(BsmtHalfBath, 0),
         BsmtUnfSF = replace_na(BsmtUnfSF, 0),
         EnclosedPorch = replace_na(EnclosedPorch, 0),
         Fireplaces = replace_na(Fireplaces, 0),
         GarageArea = replace_na(GarageArea, 0),
         GarageCars = replace_na(GarageCars, 0),
         HalfBath = replace_na(HalfBath, 0),
         KitchenAbvGr = replace_na(KitchenAbvGr, 0),
         LotFrontage = replace_na(LotFrontage, 0),
         OpenPorchSF = replace_na(OpenPorchSF, 0),
         PoolArea = replace_na(PoolArea, 0),
         ScreenPorch = replace_na(ScreenPorch, 0),
         TotRmsAbvGrd = replace_na(TotRmsAbvGrd, 0),
         WoodDeckSF = replace_na(WoodDeckSF, 0))
```

There are some numerical features that are really categories that will need to be cleaned up

```{r}
clean_MSSubClass <- function(x){
  return(paste0("SC", as.character(x)))
}

clean_MoSold <- function(x){
  if(x > 9){return(as.character(x))}
  else{
    return(paste0("0", as.character(x)))
  }
}

kaggle <- kaggle %>%
  rowwise() %>%
  mutate(MSSubClass = clean_MSSubClass(MSSubClass)) %>%
  mutate(MoSold = clean_MoSold(MoSold)) %>%
  ungroup() %>%
  mutate(MSSubClass = as.factor(MSSubClass),
         MoSold = as.factor(MoSold)) %>%
  as.data.frame()
```

@@

    d_table <- kaggle %>%
      group_by(.dots = variable) %>%
      summarise(count = n()) %>%
      kable() %>%
      kable_styling()
  } else {
    p <- ggplot(kaggle, aes_string(variable)) + 
      geom_density()
    q <- quantile(d, na.rm = TRUE)
    d_table <- data.frame(Min = min(d, na.rm = TRUE),
                          First = q[2],
                          Median = median(d, na.rm = TRUE), 
                          Mean = mean(d, na.rm = TRUE),
                          Third = q[4],
                          Max = max(d, na.rm = TRUE)) %>%
      rename("First Quartile" = First,
             "Third Quartile" = Third) %>%
      gather(Measure, Value) %>%
      kable() %>%
      kable_styling()
  }
  print(p)
  print(d_table)
}
```

@@

```{r, echo=TRUE}
correlation_matrix <- kaggle %>%
  select(TotalBsmtSF, GrLivArea, YearBuilt) %>%
  cor() %>%
  as.matrix()

correlation_matrix %>%
  kable() %>%
  kable_styling()
```

Test the hypothesis that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval.

```{r, results="asis", tidy=TRUE}
zero_vars <- c()

variables <- kaggle %>%
  select(-SalePrice) %>%
  names()

cat("<table><thead><tr><th>Variable</th><th>Correlation</th><th>Equal to Zero</th></tr></thread><tbody>")

for(variable in variables){
  d <- kaggle[,names(kaggle) == variable]
  if(is.numeric(d)){
    results <- cor.test(kaggle$SalePrice, d, conf.level = 0.8)
    if(0 > results$conf.int[1] & results$conf.int[2] > 0){
      hypothesis_test_results <- "Yes"
      zero_vars <- c(zero_vars, variable)
    } else {
      hypothesis_test_results <- "No"
    }
    cat(paste("<tr><td>",variable,"</td><td>", cor(kaggle$SalePrice, d), "</td><td>", hypothesis_test_results, "</td>"))
  }
}
cat("</tbody></table>")
```

Discuss the meaning of your analysis. Would you be worried about familywise error? Why or why not?

**There are `r length(zero_vars)` variables that have no correlation with the sale price.  I would be worried be woried about family-wise error because I just ran `r length(variable)` hypothesis tests.  The family-wise error rate would be: ${FWER} = 1 - (1 - .2)^`r length(zero_vars)` = 1 - `r .8 ^ length(zero_vars)` = `r 1 - (.8^ length(zero_vars))`$.  This means there is a high probability of committing a Type I error.**

@@

Invert your correlation matrix from above. (This is known as the precision matrix and contains variance inflation factors on the diagonal.)

```{r}
precision_matrix <- inv(correlation_matrix)

precision_matrix %>%
  kable() %>%
  kable_styling()
```

Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix.

**This will generate the identity matrix.**

```{r}
correlation_matrix %*% precision_matrix %>%
  round() %>%
  kable() %>%
  kable_styling()
```

```{r}
precision_matrix %*% correlation_matrix %>%
  round() %>%
  kable() %>%
  kable_styling()
```

Conduct LU decomposition on the matrix.

```{r}
A <- Matrix::expand(lu(correlation_matrix))

A$L %>%
  as.matrix() %>%
  kable() %>%
  kable_styling()

A$U %>%
  as.matrix() %>%
  kable() %>%
  kable_styling()
```
